Near Optimal Behavior with Approximate State Abstraction

### 2016-02-02 paper/background.tex
pass partially complete

First, some background.\dnote{Mmm. Hmmm? mmmm.}

%MDP/SDM Background and Notation
\subsection{\acp{MDP} and Sequential Decision Making}
An \ac{MDP} is a five-tuple: $\langle \mathcal{S}, \mathcal{A}, \mathcal{T}, \mathcal{R}, \gamma \rangle$: $\mathcal{S}$ is a finite state space; $\mathcal{A}$ is the set of actions available to the agent; $\mathcal{T}$ denotes $\mathcal{T}(s' \mid s,a)$, the probability of an agent applying action $a \in \mathcal{A}$ in state $s \in \mathcal{S}$ resulting in arriving in $s' \in \mathcal{S}$; $\mathcal{R}(s)$ denotes the reward received by the agent arriving in state $s$; $\gamma \in [0, 1]$ is a discount  factor that defines how much the agent prefers future rewards over immediate rewards. We assume without loss of generality that all reward functions are normalized to $[0,1]$.

The goal of an agent in an \ac{MDP} is to solve for a policy that maps states to actions: $\pi: \mathcal{S} \rightarrow \mathcal{A}$. In particular, the agent wants to solve for the policy that maximizes its expected discounted reward from any state. This goal policy is denoted $\pi^*$. We denote the expected discounted reward for following policy $\pi$ from state $s$ as the value of the state under that policy, $V^\pi(s)$. We similarly denote the expected discounted reward for taking action $a \in \mathcal{A}$ and then following policy $\pi$ from state $s$ forever after as $Q^\pi(s,a)$. Lastly, we denote the value and $Q$ functions under the optimal policy as $V^*$ and $Q^*$, respectively. \enote{Add citation -- either Littman review paper or Sutton/Barto or Puterman}

\enote{Define value of state under a policy}
\dnote{Q function, since we use it a lot}

%Lihong Section
\subsection{Abstraction Notation}
We build upon the notation used by \citeauthor{li2006towards} \enote{This should probably say more}.

We understand an abstraction as a mapping from the state space of a ground MDP $M_G$ to that of an abstract MDP $M_A$ using a state-aggregation scheme. $M_G = \langle \mathcal{S}_G, \mathcal{A}, \mathcal{T}_G, \mathcal{R}_G, \gamma \rangle$ and $M_A = \langle \mathcal{S}_A, \mathcal{A}, \mathcal{T}_A, \mathcal{R}_A, \gamma \rangle$.

\dnote{Might be worth noting that $\phi$ induces a full abstract MDP in the ``intuitively" section}

@@

% Abstract policy in ground.
In general, we are interested in how policies defined over the abstract \ac{MDP} perform in the ground \ac{MDP}. To evaluate a policy $\pi_A:\mathcal{A}\rightarrow\mathcal{A}$, we map a ground state into the abstract state space, and query the abstract policy for behavior. Specifically:
\begin{equation}
V^{\pi_A}(s) = \mathcal{R}_G(s,\pi_A(\phi(s))) + \gamma \sum_{s' \in \mathcal{S}_G} \mathcal{T}_G(s,\pi_A(\phi(s)),s')V^{\pi_A}(s').
\end{equation}

\enote{Define $V_G$ and $V_A$ (value function in ground and value function in abstract)}
\dnote{This should be done in the background section, right? Just defining a value fund under a policy}
\enote{Say more about the weighting, w(s), (that it sums to 1 but that's all we assume)}

@@

\begin{defn}
A model-irrelevance abstraction, $\phi_{model}$, is one such that:
\begin{multline}
\phi_{model}(s_1) = \phi_{model}(s_2) \rightarrow \forall a\ \mathcal{R}_G(s_1,a) = \mathcal{R}_G(s_2,a). \\
\end{multline}
and
\begin{multline}
\phi_{model}(s_1) = \phi_{model}(s_2) \rightarrow \\ \forall a \sum_{s' \in \phi^{-1}(\phi(s_1))} \mathcal{T}_G(s_1,a,s') =\sum_{s' \in \phi^{-1}(\phi(s_2))} \mathcal{T}_G(s_2,a,s'). \\
\end{multline}
\end{defn}

In words, zzz.\dnote{Insert details about each $\phi$ above}

>>>>

%First, some background.

%MDP/SDM Background and Notation
\subsection{\acp{MDP} and Sequential Decision Making}
An \ac{MDP} is a five-tuple: $\langle \mathcal{S}, \mathcal{A}, \mathcal{T}, \mathcal{R}, \gamma \rangle$: $\mathcal{S}$ is a finite state space; $\mathcal{A}$ is the set of actions available to the agent; $\mathcal{T}$ denotes $\mathcal{T}(s' \mid s,a)$, the probability of an agent transitioning to state $s' \in \mathcal{S}$ after applying action $a \in \mathcal{A}$ in state $s \in \mathcal{S}$; $\mathcal{R}(s)$ denotes the reward received by the agent arriving in state $s$; $\gamma \in [0, 1]$ is a discount  factor that defines how much the agent prefers future rewards over immediate rewards. We assume without loss of generality that all reward functions are normalized to $[0,1]$.

The goal of an agent in an \ac{MDP} is to solve for a policy that maps states to actions: $\pi: \mathcal{S} \rightarrow \mathcal{A}$. In particular, the agent wants to solve for the policy that maximizes its expected discounted reward from any state. This goal policy is denoted $\pi^*$. We denote the expected discounted reward for following policy $\pi$ from state $s$ as the value of the state under that policy, $V^\pi(s)$. We similarly denote the expected discounted reward for taking action $a \in \mathcal{A}$ and then following policy $\pi$ from state $s$ forever after as $Q^\pi(s,a)$, defined by the Bellman Equation as:
\begin{equation}
Q^\pi(s,a) = \mathcal{R}(s,a) + \gamma \sum_{s'} \mathcal{T}(s,a,s') Q^\pi(s',\pi(s))
\end{equation}
Similarly, the value function defined under a given policy, denoted $V^\pi(s)$, is defined as:
\begin{equation}
V^\pi(s) = \max_a Q^\pi(s,a)
\end{equation}

Lastly, we denote the value and $Q$ functions under the optimal policy as $V^*$ and $Q^*$, respectively~\cite{kaelbling1996reinforcement}.

%Lihong Section
\subsection{Abstraction Notation}
We build upon the notation used by \citeauthor{li2006towards}, who introduced a unifying theoretical framework for state abstraction in \acp{MDP} and sequential decision making. We build heavily on their notation, and discuss other similarities in our work in the next section.

We understand an abstraction as a mapping from the state space of a ground MDP $M_G$ to that of an abstract MDP $M_A$ using a state-aggregation scheme. Consequently, this mapping induces an abstract MDP. Let:$M_G = \langle \mathcal{S}_G, \mathcal{A}, \mathcal{T}_G, \mathcal{R}_G, \gamma \rangle$ and $M_A = \langle \mathcal{S}_A, \mathcal{A}, \mathcal{T}_A, \mathcal{R}_A, \gamma \rangle$.

@@

We refer to the weight associated with a ground state, $s \in \mathcal{S}_G$ by $\omega(s)$. The only restriction placed on the weighting schema is that it induces a probability distribution on the states:
\begin{equation}
\left(\sum_{s \in \mathcal{S}_G} \omega(s)\right) = 1,\hspace{6mm} \forall_{s \in \mathcal{S}_G} \omega(s) \in [0,1]
\end{equation}

% Abstract policy in ground.
In general, we are interested in how policies defined over the abstract \ac{MDP} perform in the ground \ac{MDP}. To evaluate a policy $\pi_A:\mathcal{A}\rightarrow\mathcal{A}$, we map a ground state into the abstract state space, and query the abstract policy for behavior. Specifically:
\begin{equation}
V^{\pi_A}(s) = \mathcal{R}_G(s,\pi_A(\phi(s))) + \gamma \sum_{s' \in \mathcal{S}_G} \mathcal{T}_G(s,\pi_A(\phi(s)),s')V^{\pi_A}(s').
\end{equation}
\dnote{I don't think this is right. Should be the max action? Or am I confused? My reading of the Q function from above is then possibly confused...}

%\enote{Define $V_G$ and $V_A$ (value function in ground and value function in abstract)}
%\dnote{This should be done in the background section, right? Just defining a value fund under a policy}

@@

\begin{defn}
A model-irrelevance abstraction, $\phi_{model}$, is one such that:
\begin{multline}
\phi_{model}(s_1) = \phi_{model}(s_2) \rightarrow \forall a\ \mathcal{R}_G(s_1,a) = \mathcal{R}_G(s_2,a).
\end{multline}
and
\begin{multline}
\phi_{model}(s_1) = \phi_{model}(s_2) \rightarrow \\ \forall a \sum_{s' \in \phi^{-1}(\phi(s_1))} \mathcal{T}_G(s_1,a,s') = \sum_{s' \in \phi^{-1}(\phi(s_2))} \mathcal{T}_G(s_2,a,s').
\end{multline}
\end{defn}

Intuitively, a $Q^*$-irrelevance is one that is guaranteed to compress states that share optimal Q functions. Critically, the direction of the implication is such that the abstraction is not guaranteed to abstract all states for which this condition is true, only that, if the abstraction function collapses two (or more) states to an abstract state, that those states have the same $Q^*$. Computing such a function is as difficult as solving the MDP, as it requires knowledge of the $Q*$, which is a sufficient condition for knowing the optimal policy. This is one of the motivations for investigating approximate abstractions, as we anticipate that an approximation of $Q^*$ is the sort of knowledge a learner can infer online.

A model-irrelevance abstraction is one that is guaranteed to compress states that transition to the same compressed states, and have identical rewards. Critically, this need not require that the ground transition functions of any two compressed states are identical; only that their corresponding abstract state transitions {\it in the abstract} MDP are identical. Unlike the $Q^*$ irrelevance, in the planning setting we anticipate knowing the full transition function and reward function up front, so it is feasible that an abstraction function of this kind may be used. In a model based learning setting, we expect an approximate abstraction akin to model-irrelevance to be more realistic, as we anticipate reinforcement learners to be able to approximate the information content of the transition function and reward function during learning.

### 2016-02-02 paper/paper.tex
pass partially complete

%THeorem declarations
\newtheorem{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem{clm}{Claim}
\newtheorem{defn}{Definition}

\icmltitlerunning{Suboptimality Bounds for Approximate State Abstraction}

@@

\twocolumn[
\icmltitle{Optimality Bounds for Approximate State Abstraction}

@@

% --- SECTION: Conclusion ---
\section{Conclusion}

% Summary

% Future Work
\begin{enumerate}
\item Learning Phi
\begin{itemize}
\item Exploration vs. Exploitation problem is different while trying to learn Phi
\end{itemize}
\item Compressibility
\begin{itemize}
\item Relationship between approximate abstract and compressibility
\end{itemize}
\item POMDP and abstraction
\end{enumerate}

>>>>

%Theorem declarations
\newtheorem{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem{clm}{Claim}
\newtheorem{defn}{Definition}

\icmltitlerunning{Near Optimal Behavior with Approximate State Abstraction}

@@

\twocolumn[
% Suboptimality Bounds for Approximate State Abstraction
\icmltitle{Near Optimal Behavior with Approximate State Abstraction}

@@

% --------------------------------------
% --- SECTION: Conclusion ---
% --------------------------------------
\input{conclusion.tex}

### 2016-02-02 paper/related_work.tex
pass partially complete

\subsection{Approximate State Abstraction}
\citeauthor*{dean1997model} investigated partitioning an MDP's state space into $\epsilon$-homogenous blocks, which are defined as clusters of states whose transition model and reward function are within $\epsilon$ of each other. They provide bounds on the error incurred as a result of using the optimal policy defined over these blocks in the original MDP. This method of abstraction is effectively characterized as a function belonging to the approximate class $\phi_{model}$.
Several approaches build on the notion of $\epsilon$-homogeneity. \citeauthor*{even2003approximate} analyzed different distance metrics used in the process of identifying $\epsilon$-homogenous partitions. \citeauthor*{ortner2013adaptive} developed an algorithm for learning partitions in an online setting by taking advantage of the confidence bounds for $T$ and $R$ provided by UCRL~\cite{auer2009near}.

>>>>

\subsection{Approximate State Abstraction}
\citeauthor*{dean1997model} investigated partitioning an \ac{MDP}'s state space into $\epsilon$-homogenous blocks, which are defined as clusters of states whose transition model and reward function are within $\epsilon$ of each other. They develop an algorithm called Interval Value Iteration (IVI) that converges to the correct bounds on a family of abstracted MDPs called a Bounded MDP. This method of abstraction is effectively characterized as a function belonging to the approximate class $\phi_{model}$.
Several approaches build on the notion of $\epsilon$-homogeneity. \citeauthor*{even2003approximate} analyzed different distance metrics used in the process of identifying $\epsilon$-homogenous partitions. \citeauthor*{ortner2013adaptive} developed an algorithm for learning partitions in an online setting by taking advantage of the confidence bounds for $T$ and $R$ provided by UCRL~\cite{auer2009near}.

### 2016-02-02 paper/theory_results.tex
pass partially complete

@@

\begin{lma}
\label{lma:Q*}
When $\ep_{Q^*}$ is used to create the abstract state space $\mathcal{S}_A$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V(s) - V^{\pi^*_{A}}(s) | \leq \frac{2\epsilon}{(1-\gamma)^2}.
\end{equation}
\end{lma}

\textbf{Proof:}
We begin by showing that Q values in the abstract \ac{MDP} closely resemble those in the ground \ac{MDP}.

@@

\begin{clm}
Q-Values in the abstract \ac{MDP} closely resemble those in the ground \ac{MDP}:
\begin{equation}
\label{eq:Q*Claim1}
\forall_{s_G \in \mathcal{S}_G, a} |Q_G(s_G, a) - Q_A(\epQ(s_G), a)| \leq \frac{\epsilon}{1-\gamma}.
\end{equation}
\end{clm}

Consider a temporally heterogeneous \ac{MDP}, $M^T = \langle \mathcal{S}_T, \mathcal{A}_G, \mathcal{R}_T, \mathcal{T}_T, \gamma \rangle$, parameterized by integer $T$, such that for the first $T$ time steps the reward function, transition dynamics and state space are those of the abstract MDP, $M_A$, and after $T$ time steps the reward function, transition dynamics and state spaces are those of the ground MDP, $M_G$. Thus,
\begin{equation}
\mathcal{S}_T = \begin{cases}
\mathcal{S}_G,& T \leq 0, \\
\mathcal{S}_A,& \text{o/w}, \\
\end{cases}
\end{equation}
\begin{equation}
\mathcal{R}_T(s,a) = \begin{cases}
\mathcal{R}_G(s,a),& T \leq 0, \\
\mathcal{R}_A(s, a),& \text{o/w}, \\
\end{cases}
\end{equation}
and
\begin{equation}
\mathcal{T}_T(s,a,s') = \begin{cases}
\mathcal{T}_G(s,a,s'),& T \leq 0, \\
\underset{{g \in X(s)}}{\sum}\left[w(g)T_G(g, a, s')\right],& T = 1, \\
\mathcal{T}_A(s,a,s'),& \text{o/w}. \\
\end{cases}
\end{equation}

>>>>

% Explanation of the above.
The first statement is relating Q values defined over the ground MDP with Q values defined over the abstract MDP, while the second statement is comparing the optimal Q values in the ground MDP, with the Q values in the ground MDP under the abstract policy. This is because our learners will solve for the optimal policy in the abstract, and use it in the ground MDP by mapping each ground state to its abstract state and querying the abstract policy for behavior. That is:
\begin{equation}
\pi_G(s) = \pi_A(\phi(s))
\end{equation}

@@

% Q^* Lemma
\begin{lma}
\label{lma:Q*}
When $\ep_{Q^*}$ is used to create the abstract state space $\mathcal{S}_A$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V(s) - V^{\pi^*_{A}}(s) | \leq \frac{2\epsilon}{(1-\gamma)^2}.
\end{equation}
\end{lma}

% Proof
\textbf{Proof:}
We begin by showing that Q values in the abstract \ac{MDP} closely resemble those in the ground \ac{MDP}.

@@

% Claim 1:
\begin{clm}
Q-Values in the abstract \ac{MDP} closely resemble those in the ground \ac{MDP}:
\begin{equation}
\label{eq:Q*Claim1}
\forall_{s_G \in \mathcal{S}_G, a} |Q_G(s_G, a) - Q_A(\epQ(s_G), a)| \leq \frac{\epsilon}{1-\gamma}.
\end{equation}
\end{clm}

Consider a temporally heterogeneous \ac{MDP}, $M^T = \langle \mathcal{S}_T, \mathcal{A}_G, \mathcal{R}_T, \mathcal{T}_T, \gamma \rangle$, parameterized by integer $T$, such that for the first $T$ time steps the reward function, transition dynamics and state space are those of the abstract MDP, $M_A$, and after $T$ time steps the reward function, transition dynamics and state spaces are those of the ground MDP, $M_G$. Thus,
\begin{equation}
\mathcal{S}_T = \begin{cases}
\mathcal{S}_G,& T \leq 0, \\
\mathcal{S}_A,& \text{o/w},
\end{cases}
\end{equation}
\begin{equation}
\mathcal{R}_T(s,a) = \begin{cases}
\mathcal{R}_G(s,a),& T \leq 0, \\
\mathcal{R}_A(s, a),& \text{o/w},
\end{cases}
\end{equation}
and
\begin{equation}
\mathcal{T}_T(s,a,s') = \begin{cases}
\mathcal{T}_G(s,a,s'),& T \leq 0, \\
\underset{{g \in X(s)}}{\sum}\left[w(g)T_G(g, a, s')\right],& T = 1, \\
\mathcal{T}_A(s,a,s'),& \text{o/w}.
\end{cases}
\end{equation}
